---
title: "Analytics Report for Ekstraklasa"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme:
      version: 4
      bg: "#101010"
      fg: "#FDF7F7" 
      primary: "#ED79F9"
      navbar-bg: "#3ADAC6"
      base_font: 
        google: Prompt
      heading_font:
        google: Sen
      code_font:
        google: 
          # arguments to sass::font_google() 
          family: JetBrains Mono
          local: false
    source_code: https://github.com/LukaszChrostowski/Football_Results_Predictions
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(plotly)
library(ggplot2)
library(dplyr)
library(lubridate)
library(lattice)
library(shiny)

load("output/processed_data.Rdata")
load("data/contingencyTable.Rdata")
df <- proccessed_data
```


```{r}
# FUNCTIONS

# Descriptive plots for the choosen season
ConvFun <- function(data) {
  
  data[data == "H"] <- 1
  data[data == "D"] <- 2
  data[data == "A"] <- 3
  
  for (col in colnames(data)) {
    
    data[, col] <- as.numeric(data[, col])
    
  }
  data <- as.matrix(data)
  data
}

levelplotFun <- function(data){
  
  plt <- levelplot(t(data[c(nrow(data):1), ]), scales=list(x=list(rot=45)),
                   col.regions = colorRampPalette(c("green", "yellow", "red"))(100),
                   colorkey=FALSE,
                   xlab = "Away Team", ylab = "Home Team")
  plt
}
```

{.sidebar}
======================================================================

```{r}
selectInput("x_variable", label = "First variable:",
            choices = c("Posiadanie piłki Gospodarz", "Sytuacje bramkowe Gospodarz", "Strzały na bramkę Gospodarz"  ,  
                        "Strzały niecelne Gospodarz", "Rzuty rożne Gospodarz", "Interwencje bramkarzy Gospodarz",
                        "Faule Gospodarz", "Rzuty wolne Gospodarz", "Auty bramkowe Gospodarz", 
                        "Posiadanie piłki Gość", "Sytuacje bramkowe Gość", "Strzały na bramkę Gość",      
                        "Strzały niecelne Gość", "Interwencje bramkarzy Gość",    
                        "Strzały zablokowane Gospodarz", "Strzały zablokowane Gość",     
                        "Ataki Gospodarz",  "Niebezpieczne ataki Gospodarz",
                        "Ataki Gość",   "Niebezpieczne ataki Gość", "Podania Gospodarz",          
                        "Podania Gość", "Podania celne Gospodarz", "Podania celne Gość"),
  selected = "`Posiadanie piłki Gospodarz`")
selectInput("y_variable", label = "Second variable:",
            choices = c("Posiadanie piłki Gospodarz", "Sytuacje bramkowe Gospodarz", "Strzały na bramkę Gospodarz"  ,  
                        "Strzały niecelne Gospodarz", "Rzuty rożne Gospodarz", "Interwencje bramkarzy Gospodarz",
                        "Faule Gospodarz", "Rzuty wolne Gospodarz", "Auty bramkowe Gospodarz", 
                        "Posiadanie piłki Gość", "Sytuacje bramkowe Gość", "Strzały na bramkę Gość",      
                        "Strzały niecelne Gość", "Interwencje bramkarzy Gość",    
                        "Strzały zablokowane Gospodarz", "Strzały zablokowane Gość",     
                        "Ataki Gospodarz",  "Niebezpieczne ataki Gospodarz",
                        "Ataki Gość",   "Niebezpieczne ataki Gość", "Podania Gospodarz",          
                        "Podania Gość", "Podania celne Gospodarz", "Podania celne Gość"),
  selected = "`Posiadanie piłki Gość`")
selectInput("club", label = "Club to Analyze",
            choices = unique(df$Gospodarz), selected = "Lech Poznań")
selectInput("season", label = "Season to Analyze",
            choices = c(as.vector(unique(df$Sezon)), "all"), selected = "all")
```

TODO DESCRIPTION / LEGEND

Author 1: [Łukasz Chrostowski](https://github.com/LukaszChrostowski) 

Author 2: [Piotr Chlebicki](https://github.com/Kertoo)

Scatter plots
======================================================================

Row {data-height=650}
-----------------------------------------------------------------------
### Scatter plots

```{r}
renderPlotly({
  plot <- ggplot(data = df, aes_string(x = paste("`", input$x_variable, "`", sep = ""), y = paste("`", input$y_variable, "`", sep = ""))) +
    geom_point() +
    xlab(input$x_variable) +
    ylab(input$y_variable) +
    ggtitle(paste(input$x_variable, "vs", input$y_variable))
  plot
  
})
```


Plots by club
======================================================================
Row
-----------------------------------------------------------------------

### With whom did the club face most often?

```{r}
renderPlotly({
  club <- input$club
  df_club <- df
    # Z kim najczęściej się mierzyli
  dff1 <- df_club %>% filter(Gospodarz == club) %>% group_by(Gość) %>% summarise(n = n()) %>% arrange(-n)
  dff2 <- df_club %>% filter(Gość == club) %>% group_by(Gospodarz) %>% summarise(n = n()) %>% arrange(-n)
  colnames(dff1) <- colnames(dff2) <- c("Club", "n")
  merge_df <- merge(dff1, dff2, by = "Club")
  merge_df$n <- merge_df$n.x + merge_df$n.y
  merge_df <- merge_df %>% arrange(-n) %>% head(8)
  
  g1 <- ggplot(data = merge_df, aes(x = factor(Club, levels = Club), y = n, fill = Club)) +
    geom_bar(stat = "identity") +
    xlab("Club") +
    ylab("Number of the games") +
    scale_fill_manual(values = 1:nrow(merge_df)) +
    guides(fill = FALSE) +
    theme_dark()
  g1
})
```

Row
-----------------------------------------------------------------------

### With whom did the club win most often?

```{r}
renderPlotly({
    club <- input$club
    df_club <- df
     # Z kim najczęściej wygrywali
    dff1 <- df_club %>% filter(Gospodarz == club) %>% group_by(Gość) %>% filter(Wynik == "H") %>% summarise(n = n()) %>% arrange(-n)
    dff2 <- df_club %>% filter(Gość == club) %>% group_by(Gospodarz) %>% filter(Wynik == "A") %>% summarise(n = n()) %>% arrange(-n)
    colnames(dff1) <- colnames(dff2) <- c("Club", "n")
    merge_df <- merge(dff1, dff2, by = "Club") 
    merge_df$n <- merge_df$n.x + merge_df$n.y
    merge_df <- merge_df %>% arrange(-n) %>% head(8)
    
    g2 <- ggplot(data = merge_df, aes(x = factor(Club, levels = Club), y = n, fill = Club)) +
      geom_bar(stat = "identity") +
      xlab("Club") +
      ylab("Number of the games") +
      scale_fill_manual(values = 1:nrow(merge_df)) +
      guides(fill = FALSE) +
      theme_dark()
    g2
    
})
```

Row
-----------------------------------------------------------------------

### With whom did the club draw most often?

```{r}
renderPlotly({
    club <- input$club
    df_club <- df
    # Z kim najczęściej remisowali
    dff1 <- df_club %>% filter(Gospodarz == club) %>% group_by(Gość)  %>% filter(Wynik == "D") %>% summarise(n = n()) %>% arrange(-n)
    dff2 <- df_club %>% filter(Gość == club) %>% group_by(Gospodarz) %>% filter(Wynik == "D") %>% summarise(n = n()) %>% arrange(-n)
    colnames(dff1) <- colnames(dff2) <- c("Club", "n")
    merge_df <- merge(dff1, dff2, by = "Club") 
    merge_df$n <- merge_df$n.x + merge_df$n.y
    merge_df <- merge_df %>% arrange(-n) %>% head(8)
    
    g3 <- ggplot(data = merge_df, aes(x = factor(Club, levels = Club), y = n, fill = Club)) +
      geom_bar(stat = "identity") +
      xlab("Club") +
      ylab("Number of the games") +
      scale_fill_manual(values = 1:nrow(merge_df)) +
      guides(fill = FALSE) +
      theme_dark()
    g3
})
```

Row
-----------------------------------------------------------------------
### With whom did the club lose most often?

```{r}
renderPlotly({
    club <- input$club
  df_club <- df
  # Z kim najczęściej przegrywali
  dff1 <- df_club %>% filter(Gospodarz == club) %>% group_by(Gość) %>% filter(Wynik == "A") %>% summarise(n = n()) %>% arrange(-n)
  dff2 <- df_club %>% filter(Gość == club) %>% group_by(Gospodarz) %>% filter(Wynik == "H") %>% summarise(n = n()) %>% arrange(-n)
  colnames(dff1) <- colnames(dff2) <- c("Club", "n")
  merge_df <- merge(dff1, dff2, by = "Club")
  merge_df$n <- merge_df$n.x + merge_df$n.y
  merge_df <- merge_df %>% arrange(-n) %>% head(8)
  
  g4 <- ggplot(data = merge_df, aes(x = factor(Club, levels = Club), y = n, fill = Club)) +
    geom_bar(stat = "identity") +
    xlab("Club") +
    ylab("Number of the games") +
    scale_fill_manual(values = 1:nrow(merge_df)) +
    guides(fill = FALSE) +
    theme_dark()
  g4
})
```

Plots by season
======================================================================

Row {data-height=650}
-----------------------------------------------------------------------

### Heatmap for the results

```{r}
renderPlot({
  season <- input$season
  if (season != "all") {
    df <- df %>% filter(Sezon == season)
    ct <-  switch(season,
              "2012/13" = ContTable_1213,
              "2013/14" = ContTable_1314,
              "2014/15" = ContTable_1415,
              "2015/16" = ContTable_1516,
              "2016/17" = ContTable_1617,
              "2017/18" = ContTable_1718,
              "2018/19" = ContTable_1819,
              "2019/20" = ContTable_1920,
              "2020/21" = ContTable_2021,
              "2021/22" = ContTable_2122)
  ct <- ConvFun(ct)
  levelplotFun(ct)
  } else {
  levelplotFun(ConvFun(ContTable_2122))
  }
})
```


Row
-----------------------------------------------------------------------

### Results distribution

```{r}
renderPlotly({
  
  season <- input$season
  if (season != "all") df <- df %>% filter(Sezon == season)
  
  ggplot(data = df, aes(x = Wynik, fill = Wynik)) +
    geom_bar() +
    ylab("Number of the results") +
     xlab("Result") +
    scale_fill_manual(values = unique(df$Wynik)) +
    guides(fill = FALSE)

})
```

Row
-----------------------------------------------------------------------
### Distribution of the goals

```{r}
renderPlotly({
  season <- input$season
  if (season != "all") df <- df %>% filter(Sezon == season)
  
  ggplot(data = df, aes(x = `Gole Gospodarz` + `Gole Gość` )) + 
  geom_histogram(bins = 10, color = "black", fill = "white") +
  xlab("Number of goals in the game")
})
```

Row
-----------------------------------------------------------------------

### Sum of goals during the seasons

```{r}
renderPlotly({
  season <- input$season
  if (season != "all") df <- df %>% filter(Sezon == season)
  
  goals_by_season <- df %>% group_by(Sezon) %>% summarise(sum = sum(`Gole Gospodarz` + `Gole Gość`))
  
   ggplot(data = goals_by_season, aes(x = Sezon, y = sum, fill = Sezon)) +
    geom_bar(stat = "identity") +
    xlab("Season") +
    ylab("Goals total") +
    scale_fill_manual(values = 1:nrow(goals_by_season)) +
    guides(fill = FALSE)
})
```

Row
-----------------------------------------------------------------------
### Ranking of the most effective clubs

```{r}
renderPlotly({
  season <- input$season
  if (season != "all") df <- df %>% filter(Sezon == season)
  
  # Number of goals by each club (sum, mean)
  A <- df %>% group_by(Gospodarz) %>% summarise(sum = sum(`Gole Gospodarz`)) %>% arrange(-sum)
  B <- df %>% group_by(Gość) %>% summarise(sum = sum(`Gole Gość`)) %>% arrange(-sum)
  colnames(A) <- colnames(B) <- c("Klub", "gole")
  mergeAB <- merge(A, B, by = "Klub")
  mergeAB$gole <- mergeAB$gole.x + mergeAB$gole.y
  goals_by_club <- mergeAB %>% arrange(-gole) %>% head(8)
  
  ggplot(data = goals_by_club, aes(x = factor(Klub, levels = Klub), y = gole, fill = Klub)) +
    geom_bar(stat = "identity") +
    xlab("Season") +
    ylab("Scored goals total") +
    scale_fill_manual(values = 1:nrow(goals_by_club)) +
    guides(fill = FALSE) +
    theme(axis.text.x = element_text(angle = 30, hjust = 1))
})
```

Row
-----------------------------------------------------------------------
### Ranking of the clubs by number of the red cards

```{r}
renderPlotly({
  season <- input$season
  if (season != "all") df <- df %>% filter(Sezon == season)
  
  A <- df %>% group_by(Gospodarz) %>% summarise(sum_red = sum(`Czerwone kartki Gospodarz`),
                                                sum_yellow = sum(`Żółte kartki Gospodarz`))
  B <- df %>% group_by(Gość) %>% summarise(sum_red = sum(`Czerwone kartki Gość`),
                                           sum_yellow = sum(`Żółte kartki Gość`)) 
  colnames(A) <- colnames(B) <-  c("Club", "red_cards", "yellow_cards")
  mergeAB <- merge(A, B, by = "Club")
  mergeAB$red_cards <- mergeAB$red_cards.x + mergeAB$red_cards.y
  mergeAB$yellow_cards <- mergeAB$yellow_cards.x + mergeAB$yellow_cards.y
  red_cards_by_club <- mergeAB %>% arrange(-red_cards) %>% head(8)
  
   ggplot(data = red_cards_by_club, aes(x = factor(Club, levels = Club), y = red_cards, fill = Club)) +
        geom_bar(stat = "identity") +
        xlab("Season") +
        ylab("Red cards total") +
        scale_fill_manual(values = 1:nrow(red_cards_by_club)) +
        guides(fill = FALSE) +
        theme(axis.text.x = element_text(angle = 30, hjust = 1))

})
```

Row
-----------------------------------------------------------------------

### Ranking of the clubs by number of the yellow cards
```{r}
renderPlotly({

  season <- input$season
  if (season != "all") df <- df %>% filter(Sezon == season)
  
  A <- df %>% group_by(Gospodarz) %>% summarise(sum_red = sum(`Czerwone kartki Gospodarz`),
                                                sum_yellow = sum(`Żółte kartki Gospodarz`))
  B <- df %>% group_by(Gość) %>% summarise(sum_red = sum(`Czerwone kartki Gość`),
                                           sum_yellow = sum(`Żółte kartki Gość`)) 
  colnames(A) <- colnames(B) <-  c("Club", "red_cards", "yellow_cards")
  mergeAB <- merge(A, B, by = "Club")
  mergeAB$red_cards <- mergeAB$red_cards.x + mergeAB$red_cards.y
  mergeAB$yellow_cards <- mergeAB$yellow_cards.x + mergeAB$yellow_cards.y
  red_cards_by_club <- mergeAB %>% arrange(-red_cards) %>% head(8)
  yellow_cards_by_club <- mergeAB %>% arrange(-yellow_cards) %>% head(8)
  
  g8 <- ggplot(data = yellow_cards_by_club, aes(x = factor(Club, levels = Club), y = yellow_cards, fill = Club)) +
    geom_bar(stat = "identity") +
    xlab("Season") +
    ylab("Yellow cards total") +
    scale_fill_manual(values = 1:nrow(yellow_cards_by_club)) +
    guides(fill = FALSE)
})
```

Row
-----------------------------------------------------------------------
### Number of the games per month
```{r}
renderPlotly({
  season <- input$season
  if (season != "all") df <- df %>% filter(Sezon == season)
  
  # Plots by date of the match
  df_date <- df %>% group_by(month(Data)) %>% summarise(n = n()) %>% arrange(-n) %>% head()
  colnames(df_date) <- c("month", "n")
  df_date$month <- as.factor(month.name[df_date$month])
  
  ggplot(data = df_date, aes(x = factor(month, levels = month), y = n, fill = month)) +
    geom_bar(stat = "identity") +
    xlab("Month") +
    ylab("Number of the games") +
    scale_fill_manual(values = 1:nrow(df_date)) +
    guides(fill = FALSE)
})
```

Row
-----------------------------------------------------------------------
### Number of the games per hour
```{r}
renderPlotly({
  season <- input$season
  if (season != "all") df <- df %>% filter(Sezon == season)
  
    df_date <- df %>% group_by(hour(Data)) %>% summarise(n = n()) %>% arrange(-n) %>% head()
  colnames(df_date) <- c("hour", "n")
  df_date$hour <- as.factor(df_date$hour)
  
  ggplot(data = df_date, aes(x = factor(hour, levels = hour), y = n, fill = hour)) +
    geom_bar(stat = "identity") +
    xlab("Hour") +
    ylab("Number of the games") +
    scale_fill_manual(values = 1:nrow(df_date)) +
    guides(fill = FALSE)
})
```


